|

### 2025-07-10 output/IMPS-2025-slideshow.qmd
Improve speaker notes to cut down clutter

::: notes
The Multidimensional 2PL model is one of many extensions of the 2PL model, which in this case is compensatory.

@@

::: notes
The problem with the M2PL, and other multidimensional extensions, is that its parameters do not have a clear interpretation in IRT terms, and they cannot be reparameterized.

@@

He do it in this 1985 paper, and in a subsequent one in 1991...
:::

@@

Note also that the discrimination can in turn be used to define the item difficulty, as happens in the unidimensional model.

And also, that the difficulty has two components: a "signed distance" from the origin, and a direction.

@@

It is easier to appreciate this with the graphical representation, that was proposed by Ackerman a few years later, where we can see that the item vector is applied at a point given by the difficulty, and has a length equal to the discrimination.
Note that the item direction always goes through the origin, and that the sign of the distance (positive in this case) will determine whether the vector is displaced forward or backward along that direction.
:::

@@

::: notes
However, there are two assumptions that Reckase & McKinley make when obtaining the definition of the multidimensional parameters.

The first is that the probability of answering an item correctly always increases monotonically with respect to each dimension.

The second, that constrain that you see there, implies that the item is represented in an orthogonal basis, which is why the discrimination is defined by a "general Pythagorean theorem".

So, our purpose in this work is **to relax these two assumptions**.
:::

@@

Nevertheless, with this caveat about null discrimination components, this assumption would be realistic in a "cognitive testing context": The more you know, the more likely you are to give a correct answer.
:::

>>>>

::: notes
The Multidimensional 2PL model is one of many extensions of the 2PL model.

@@

::: notes
The problem with the M2PL is that its parameters do not have a clear interpretation in IRT terms.

@@

He does it in this 1985 paper, and in a subsequent one in 1991...
:::

@@

This discrimination can be used to define the item difficulty, finding a parameterization that parallels the unidimensional 2PL model.

Note though that, in this case, the difficulty has two components: a "signed distance" from the origin, and a direction.

@@

It is easier to appreciate this with the graphical representation, that was proposed by Ackerman a few years later, where we can see that the item vector is applied at a point given by the difficulty (with its distance and direction), and has a length equal to the discrimination.
The vector direction always goes through the origin, and the sign of the distance (positive in this case), which is the opposite of the intercept sign, will determine whether the vector is displaced forward or backward along that direction.
:::

@@

::: notes
However, there are two assumptions for defining the multidimensional parameters.

First, that the probability of a correct answer increases monotonically with each dimension.

And second, this constrain you see there, which implies that the item is represented in an orthogonal basis. This is necessary for applying the "general Pythagorean theorem" as we saw before.

:::

@@

Nevertheless, with this caveat about null discrimination components, this assumption would be realistic in a "cognitive testing context": The more you know, the more likely you are to give a correct answer.
These assumptions derive from the traditional application of the model to the cognitive domain, but we want to apply multidimensional IRT to the non-cognitive domain as well, e.g. to personality measures.
So, our purpose is **to relax these two assumptions**.
:::
